---
title: "SURP_R_Workshop_Day3"
author: "Tess Cherlin"
date: "2023-07-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# This is how we would make a line graph with error bars in R
```{r}
if (!requireNamespace("ggplot2", quietly = TRUE)) {
  # If not installed, install the package
  install.packages(package_name)
}
library(ggplot2)

# Example data frame
# Replace this with your actual data
data <- data.frame(
  x = c(1, 2, 3, 4, 5),
  y = c(10, 15, 20, 18, 22),
  sd = c(1, 2, 1.5, 3, 2.5)
)

head(data)

# Create the line graph with SD bands
ggplot(data, aes(x = x, y = y)) +
  geom_line() +
  geom_ribbon(aes(ymin = y - sd, ymax = y + sd), alpha = 0.3) +
  labs(title = "Line Graph with SD Bands",
       x = "X-axis Label",
       y = "Y-axis Label")
```

# Here is how we would add multiple lines to the same graph
```{r}
# Load the ggplot2 library
if (!requireNamespace("ggplot2", quietly = TRUE)) {
  # If not installed, install the package
  install.packages(package_name)
}
library(ggplot2)

# Example data frame for line 1
data1 <- data.frame(
  x = c(1, 2, 3, 4, 5),
  y = c(10, 15, 20, 18, 22),
  sd = c(1, 2, 1.5, 3, 2.5)
)

head(data1)

# Example data frame for line 2
data2 <- data.frame(
  x = c(1, 2, 3, 4, 5),
  y = c(8, 12, 18, 15, 20),
  sd = c(1.2, 1.8, 2, 2.5, 2.2)
)

head(data2)

# Create the line graph with multiple lines
ggplot() +
  geom_line(data = data1, aes(x = x, y = y), color = "blue") +
  geom_ribbon(data = data1, aes(x = x, ymin = y - sd, ymax = y + sd), alpha = 0.3, fill = "blue") +
  geom_line(data = data2, aes(x = x, y = y), color = "red") +
  geom_ribbon(data = data2, aes(x = x, ymin = y - sd, ymax = y + sd), alpha = 0.3, fill = "red") +
  labs(title = "Line Graph with Multiple Lines",
       x = "X-axis Label",
       y = "Y-axis Label")
```
# Lets get comfortable with Katherine's data
```{r}
kData = read.csv("KatherineData.csv")

dim(kData)
head(kData)

```
# What do we need to make our line graph?
# Calculate average and sd
# Wait we did this last time!!!
```{r}
# Remove the sample column from the data
kData2 = kData
kData2$Sample <- NULL

# Make a dataframe
kData3 = data.frame(Day = c(1,2,3),
                    Control_means = colMeans(kData2[1:3,1:3]),
                    Control_sds = apply(kData2[1:3,1:3], 2, sd),
                    Exp_2.5ul_means = colMeans(kData2[4:6,1:3]),
                    Exp_2.5ul_sds = apply(kData2[4:6,1:3], 2, sd),
                    Exp_5ul_means = colMeans(kData2[7:9,1:3]),
                    Exp_5ul_sds = apply(kData2[7:9,1:3], 2, sd),
                    Exp_10ul_means = colMeans(kData2[10:12,1:3]),
                    Exp_10ul_sds = apply(kData2[10:12,1:3], 2, sd))

```

# Now let's make a single line graph with this data
```{r}
new_labels = c("Day0", " ", "Day3", " ", "Day6")

# Create the line graph with SD bands
ggplot(kData3, aes(x = Day, y = Control_means)) +
  geom_line() +
  #scale_x_discrete(breaks = c(1.0, 1.5, 2.0, 2.5, 3.0), labels =  new_labels) +
  geom_ribbon(aes(ymin = Control_means - Control_sds, ymax = Control_means + Control_sds), alpha = 0.3) +
  labs(title = "Line Graph with SD Bands",
       x = "X-axis Label",
       y = "Y-axis Label")
```

# Here is an example of how to make barplots with error bars in with ggplot2
```{r}
# Load the ggplot2 library
if (!requireNamespace("ggplot2", quietly = TRUE)) {
  # If not installed, install the package
  install.packages(package_name)
}
library(ggplot2)

# Example data frame
# Replace this with your actual data
data <- data.frame(
  group = c("A", "B", "C", "D"),
  mean = c(10, 15, 20, 18),
  sd = c(1, 2, 1.5, 3)
)

head(data)

# Create the barplot with SD error bars
ggplot(data, aes(x = group, y = mean, fill = group)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0.2, position = position_dodge(width = 0.9)) +
  labs(title = "Barplot with SD Error Bars",
       x = "Groups",
       y = "Mean Values")
```

# Plot mutliple plots in one pane
```{r}
if (!requireNamespace("patchwork", quietly = TRUE)) {
  # If not installed, install the package
  install.packages(package_name)
}
library(patchwork)

plot1 = ggplot(kData3, aes(x = Day, y = Control_means)) +
  geom_line() +
  #scale_x_discrete(breaks = c(1.0, 1.5, 2.0, 2.5, 3.0), labels =  new_labels) +
  geom_ribbon(aes(ymin = Control_means - Control_sds, ymax = Control_means + Control_sds), alpha = 0.3) +
  labs(title = "Line Graph with SD Bands",
       x = "X-axis Label",
       y = "Y-axis Label")

plot2 = ggplot(data, aes(x = group, y = mean, fill = group)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0.2, position = position_dodge(width = 0.9)) +
  labs(title = "Barplot with SD Error Bars",
       x = "Groups",
       y = "Mean Values")

plot1 + plot2
```

# Let's make a barplot with this sample data
```{r}
# Read in data
geneData = read.csv("MoreGenes_FPKM.csv")

# Get dimensions
dim(geneData)

# Look at the data
head(geneData)


# Make new dataset and calculate average and standard defiation for 4 genes



# Make barplot for 4 genes
```
# Combine as many plots as you want together into one final figure!
```{r}
if (!requireNamespace("patchwork", quietly = TRUE)) {
  # If not installed, install the package
  install.packages(package_name)
}
library(patchwork)

# Make your plots here
```

# Woohoo!!! 